---
title: "Project 1 (Reproducible Research)"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(knitr)
library(dplyr)
library(ggplot2)

## setting up workind directory.


setwd('C:/Users/choimyon/Box Sync/Clara/Prof Develp/Coursera/Reproducible/W2')


## bringing data to Rstudio.

```{r}
data <- read.csv("activity.csv")
```

## **What is mean total number of steps taken per day?**

```{r}
## dropping missing values in 'steps' column.
## calculating the total number of steps taken per day with grouping data by date.


actdata <- data %>%
	filter(!is.na(steps)) %>%
	group_by(date) %>%
	summarise(totstepaday=sum(steps))


## make a histogram of the total number of steps taken per day

```{r}
hist(actdata$totstepaday, main="Historgram of total number steps taken per day",
  xlab="Total number of steps per day")


## calculate and report the mean and median of the total number 
## of steps taken per day

summary(actdata)

## make a time series plot of a 5-minute interval (x) and the average number
## of steps taken, averaged across all days (y)

actdata_int <- aggregate(steps~interval, data, mean)

with(actdata_int, plot(steps~interval, type = "l", main="Average number of steps taken over all days",
ylab="Average number of steps"))

## Which 5-minute interval, on average across all the days in the dataset,
## contains the maximum number of steps?

actdata_int[which.max(actdata_int$steps),]
```


## **Imputing missing values**

```{r}
## total number of missing values in the dataset

sum(is.na(data))


## replace NA values for steps and interval with their column mean

data$steps[is.na(data$steps)] <- round(mean(data$steps, na.rm=TRUE))
data$interval[is.na(data$interval)] <- round(mean(data$interval, na.rm=TRUE))

head(data)

## make a histogram of the total number of steps taken each day with missing-value-imputed dataset.

actdata_fillin <- data %>%
	group_by(date) %>%
	summarise(totstepday=sum(steps))

hist(actdata_fillin$totstepday, main="Total number of steps taken per day (imputed dataset)",
xlab="Total number of steps per day")
	
## getting mean and median values of total number of steps taken per day with the fill-in dataset

summary(actdata_fillin)

## There are differences for mean and median values between missing-value-removed dataset and 
## missing-value-imputed one. Both mean and median values for mv-removed dataset is higher by 14 and 109
## respectively compared to the ones of the mv-treated dataset.
```


## **Are there differences in activity patterns between weekdays and weekends?**

```{r}
## create weekday variable indicating whether a given date is a weekday or weekend day.

data$weekday <- ifelse(weekdays(as.Date(actdata_fillin$date)) %in%
c("Saturday", "Sunday"), "weekend", "weekday")

## Create a panel plot conting a timeseries of the 5-minute interval (x) and the average number of
## steps taken, averaged acorss all weekday days and weekend days (y).

data_wk <- aggregate(steps~interval + weekday, data, mean)

qplot(interval, steps, 
      data = data_wk,
      geom=c("line"),
      ylab = "Number of steps") +
 	facet_wrap(~ weekday, ncol = 1)
